---
title: "Assignment 7"
author: "Derek Rouch"
date: "9/29/2019"
output: html_document
---

### Setting up R

To begin, I set my working directory to my GitHub repository folder. I also cleared my Global Environment and retrieved the necessary libraries. 

```{r setup}
## knitr
knitr::opts_chunk$set(echo = TRUE)

## Set working director
setwd("~/Documents/GitHub/data-science")

## Clear environment
rm(list=ls())

## Get libraries
library(tidyverse)
library(dplyr)
library(stats)
library(flexclust)
library(ggplot2)
library(LICORS)
library(knitr)
library(modelr)
library(readxl)
```

### 1. Download the student evaluation dataset from the website.

```{r stud_eval}
## Load data
stud_eval<-read_xlsx("/Users/derekrouch/Documents/GitHub/data-science/turkiyestudentevaluation_generic.xlsx")
```

### 2. Select a set of 10 related variables that you think might serve as the basis for clustering.

The variables I selected were:

**attendance:** Code of the level of attendance; values from {0, 1, 2, 3, 4}
**difficulty:** Level of difficulty of the course as perceived by the student; values taken from {1,2,3,4,5}
**Q2:** The course aims and objectives were clearly stated at the beginning of the period.
**Q3:** The course was worth the amount of credit assigned to it.
**Q5:** The class discussions, homework assignments, applications and studies were satisfactory.
**Q8:** The quizzes, assignments, projects and exams contributed to helping the learning.
**Q9:** I greatly enjoyed the class and was eager to actively participate during the lectures.
**Q11:** The course was relevant and beneficial to my professional development.
**Q13:** The Instructor's knowledge was relevant and up to date.
**Q20:** The Instructor explained the course and was eager to be helpful to students.

```{r}
## Select 10 variables
stud_eval <- stud_eval %>% select(attendance, difficulty, Q2, Q3, Q5, Q8, Q9, Q11, Q13, Q20)
```

### 3. Run the clustering algorithm `kmeanspp` to identify 3 clusters.

I ran `kmeansapp`, specifying that it shnuld use 3 clusters.

```{r cluster}
## Run cluster analysis until it converges on a stable solution
c <- kmeanspp(stud_eval, k = 3, start = "random", iter.max = 1000, nstart = 50)
table(c$cluster)
```

### 4. Summarize the clusters using the variables you selected.

I explored the clusters first by grouping them by **attendance**...

```{r summary_attendance}
## Add predictions
stud_eval$cluster <- c$cluster

## Examine relationship of cluster assignment to chars.
mycols <- c("Group 1", "Group 2", "Group 3")

## Attendance
pt <- prop.table(table(stud_eval$cluster, stud_eval$attendance), margin = 1)

## Format table
rownames(pt) <- mycols
pt <- pt * 100
pt <- round(pt, 1)

## Group by attendance
kable(pt, row.names = TRUE, col.names = c("0", "1", "2", "3", "4"))
```

...and then by **difficulty**

```{r summary_difficulty}
## Difficulty
pt2 <- prop.table(table(stud_eval$cluster, stud_eval$difficulty), margin = 1)

## Format table
rownames(pt2) <- mycols
pt2 <- pt2 * 100
pt2 <- round(pt2, 1)

## Group by difficulty
kable(pt2, row.names = TRUE, col.names = c("1", "2", "3", "4", "5"))
```

